---
title: "Reproducible Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research Week 2 Assignment
# Activity monitoring data analysis
### Author:  Fred Jensen
### Date:  April 15, 2018
## 1. Code for reading in the dataset and/or processing the data
```{r}
	activity_data <- read.csv("C:/Users/TheJensen5/Documents/R/download/Replicated Results - Week 2/activity.csv")
	activity_data$date <- as.Date(activity_data$date,"%Y-%m-%d")
	head(activity_data)
```
## 2. Histogram of the total number of steps taken each day	
### 	Determine the mean total number of steps taken per day
```{r}
	activity_data_per_day <- aggregate(activity_data$steps, by = list(activity_data$date), FUN = sum, na.rm = TRUE)
	## Rename columns of activity_data_per_day
	names(activity_data_per_day)[names(activity_data_per_day)=="Group.1"] <- "date"
	names(activity_data_per_day)[names(activity_data_per_day)=="x"] <- "steps"
	summary(activity_data_per_day)
```
###  	Create Histogram graph
``` {r}
	hist(subset(activity_data_per_day$steps, !is.na(activity_data_per_day$steps)), xlab = "Total daily Steps", main="Histogram of Total Steps by day")
```
  
###   Mean and median number of steps taken each day
```{r}
	## Calculate and report the mean and median of the total number of steps taken per day
	mean(activity_data_per_day$steps)
	median(activity_data_per_day$steps)
```
## 3. Time series plot of the average numbers of steps taken
### Determine the average daily activity pattern
```{r}
	activity_data_per_interval <- aggregate(activity_data$steps, by = list(activity_data$interval), FUN = mean, na.rm = TRUE)
	## Rename columns of activity_data_per_interval
	names(activity_data_per_interval)[names(activity_data_per_interval)=="Group.1"] <- "interval"
	names(activity_data_per_interval)[names(activity_data_per_interval)=="x"] <- "steps"
	summary(activity_data_per_interval)
```
### Plot time series of 5 minute intervals
``` {r}
	plot(activity_data_per_interval$interval,activity_data_per_interval$steps, xlab = "Total Interval Steps", main = "Times series plot of Total Steps by Interval", type ="l")
```
  
## 4. Determine which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r}
	subset(activity_data_per_interval, activity_data_per_interval$steps == max(activity_data_per_interval$steps))[1,1]
```
## 5. Code to describe and show a strategy for imputing missing data
### Impute missing values
#### Imputed values were equal to the average numbers of steps taken during each 5 minute interval
##### Calculate the total number of missing values in the dataset
``` {r}
  activity_data_imput <- activity_data
  ## calculate total number of missing values
  missing_values <- nrow(subset(activity_data_imput, is.na(steps)))
  missing_values
	## replace missing values with the mean value of the available data by interval
	activity_data_imput$steps <- with(activity_data_imput, ave(steps, interval, FUN = function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))))
	head(activity_data_imput)
	summary(activity_data_imput)
```
## 6. Histogram of the total number of steps taken each day after missing values are imputed
### 	Determine the mean total number of steps taken per day using imputed values in data
``` {r}
  activity_data_per_day_imput <- aggregate(activity_data_imput$steps, by = list(activity_data_imput$date), FUN = sum, na.rm = TRUE)
	## Rename columns of activity_data_per_day_imput
	names(activity_data_per_day_imput)[names(activity_data_per_day_imput)=="Group.1"] <- "date"
	names(activity_data_per_day_imput)[names(activity_data_per_day_imput)=="x"] <- "steps"
	summary(activity_data_per_day_imput)
```
###  	Create Histogram graph
``` {r}
	hist(subset(activity_data_per_day_imput$steps,!is.na(activity_data_per_day_imput$steps)), xlab = "Total daily Steps", main="Histogram of Total Steps by day")
```
###   Mean and median number of steps taken each day including imputed values
```{r}
	## Calculate and report the mean total number of steps taken per day including imputed values
	mean(activity_data_per_day_imput$steps)
  median(activity_data_per_day_imput$steps)
  ## compare mean and median from original data file
  mean(activity_data_per_day_imput$steps) - mean(activity_data_per_day$steps)
	median(activity_data_per_day_imput$steps) - median(activity_data_per_day$steps)
```  
## 7. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
###	Calculate and report the mean and median of the total number of steps taken per day

### Assign days to either weekday or weekend
``` {r}
  activity_data_type_of_day <- activity_data
	activity_data_type_of_day$day <- weekdays(activity_data_type_of_day$date)
	## create table that defines days of the week as either "Weekend" or "Weekday"
	day <- c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday")
	type <- c("Weekend","Weekday","Weekday","Weekday","Weekday","Weekday","Weekend")
	type_of_day <- data.frame(day, type)
	## merge activity_data_type_of_day and type_of_day
	activity_data_type_of_day <- merge(activity_data_type_of_day, type_of_day, by.Exposuredata_consol = "day", by.Withdata_consol = "day", all.x = T)
	head(activity_data_type_of_day)
```
###aggregate by interval and type
``` {r}
  activity_data_type_of_day_interval <- aggregate(activity_data_type_of_day$steps, by = list(activity_data_type_of_day$type, activity_data_type_of_day$interval), FUN = sum, na.rm = TRUE)
  	## Rename columns of activity_data_per_day_imput
	names(activity_data_type_of_day_interval)[names(activity_data_type_of_day_interval)=="Group.1"] <- "type"
	names(activity_data_type_of_day_interval)[names(activity_data_type_of_day_interval)=="Group.2"] <- "interval"
	names(activity_data_type_of_day_interval)[names(activity_data_type_of_day_interval)=="x"] <- "steps"
  head(activity_data_type_of_day_interval)
```
### Create the panel plot graph
``` {r}
library(ggplot2)
ggplot(activity_data_type_of_day_interval, aes(x=interval, y = steps, colour = type)) + geom_line() +
	facet_grid(type ~.) + xlab("Interval") + ylab("Mean of Steps") +
	ggtitle("Comparison of Average Steps in Each Interval")
```